Application of Micro-Rotors to Induce Crystal Nucleation: Challenges, Takeaways, and Future Prospects\\ PhD Thesis

### 2025-01-26 Latex_files/Thesis/chapters/chapter 4/main.tex
updates

\chapter{Complex Langevin dynamics of spherical dimers}
\label{chapter:langevin_dynamics}
Much of the calibration theory discussed in Chapter 2 
assumes that the target particle in question is a single 
sphere, one whose scattering and motion is easily computed. 
However, while working with dense colloidal suspensions, 
one often ends up trapping more than one sphere. Li and 
Arlt \cite{Li2008} studied the case of two microspheres 
trapped in a single OT and found that multiple trapped 
beads could be mistaken for a single trapped bead with 
altered trap stiffness (see \ref{sec:harmonic_traps}). 
Theoretical studies on the case of two trapped microspheres 
by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics 
based model to show that the two trapped beads are brought 
into physical contact with each other by optical forces 
and they also calculated the axial equilibrium positions 
of the two trapped beads as a function of their size. 
Experiments in \cite{Praveen2016} confirmed that the two 
trapped beads indeed experience different trap stiffnesses 
in the vicinity of the same potential well.

There are further discussions looking into the dynamics of 
a whole host of asymmetrically shaped particles \cite{Loudet2014, 
ShengHua2005, Chetana2022}, their results all showing that 
predicting the behaviour an arbitrary shaped particle comes 
with great difficulty due to the fact that the optical force 
is dependent on a greater number of variables such as 
orientation and size factors.

With the initial goal of the project being to induce 
nucleation events via a rotating sphere, the aim of this 
chapter is to - in a limited capacity simulate and investigate 
the influence of a second particle being bound to our target 
sphere. The choice of a dimer, instead of an amorphous solid 
that might better represent a growing crystalline solid, 
allows us to consider how the dynamics of the aggregate change 
by varying a single parameter, namely the size ratio of the 
two spheres. Additionally, attempting to simulate an amorphous 
aggregate is rather difficult as calculating the optical force 
and torque is computationally slow. We build upon the works of 
Vigilante \textit{et al} \cite{Vigilante2020} to consider 
asymmetric spherical dimers and how varying size parameters 
alters the dynamics and additionally makes characterising their 
interactions within an optical trap more cumbersome.

In this chapter we will consider how the addition of a 
second sphere changes the trapping dynamics by introducing 
multiple equilibrium positions. These positions are 
dependent on both the size ratio of the dimer and its 
orientation relative to the trap. This is important to 
address, for while the the dimers' diffusive behaviour 
is the same, regardless of position and orientation, 
how it interacts with the trap may produce the wildly 
different results. Furthermore, we look at how our dimers 
interact with circularly polarised light, this is especially 
pertinent as the original plan was to utilise circularly 
polarised light to generate fluid flow in a supersaturated 
solution. The rotational motion of dimers in circularly 
polarised light goes against much of the established 
literature and suggests that the transfer of spin angular 
momentum is enabled by combining particles together. And 
lastly, we demonstrate how a quadrant photo diode, performs 
in characterising the actual interactions between a dimer 
and the optical trap. We do so to demonstrate that much 
of the dimer's trajectory information is either lost or 
poorly described by simple calibration techniques, making 
their motion difficult to characterise.

\subsection{Simulation Parameters}
As a paradigmatic example, consider a dimer suspended in 
water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
a Gaussian beam (more specifically a Laguerre-Gaussian 
beam of mode $[0,0]$, see \eqref{eq:incident}), the beam 
is focused by a objective with numerical aperture of 1.2 
and is x polarised. The size ratio of the two spheres is 
given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ 
\mu m$ unless specified otherwise; the dimer's orientation 
is given by a unit vector connecting the centres of both 
spheres (see \ref{sec:sim_parameters}). We define the 
'standard' orientation as being aligned with the direction 
of beam propagation direction - and therefore the 'inverted' 
orientation is defined when the dimer is orientated against 
the direction of beam orientation (see left hand side of 
figure~\ref{fig:paradigmatic}).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Positional and Orientational dependence of Trapping forces}
\label{sec:eq_positions}
If we wanted to start from first principles and determine 
the trap strength on our target particle the first step 
would be to locate the equilibrium position relative to 
the trap focus. For a single sphere it is easy to enough 
to understand that its centre of mass will be drawn to 
focal point of the laser due to gradient forces, once there 
the force is analogous to a harmonic spring with a fixed 
trap stiffness (see fig.\ref{fig:harmonic_trap}). Now, if 
we consider instead a dimer, we now have two spheres both 
being drawn to the focus along by the same gradient force; 
in addition the scattering force is significantly more complex 
due to both spheres scattering the electromagnetic fields. 
This mutual scattering between individual spheres is what 
makes simulating spherical aggregates far more difficult 
compared to a single sphere, and even harder still to 
predict the position where the dimer's centre of mass is 
in equilibrium.

@@

We can see that both equilibrium positions have comparable 
axial trap stiffness ($\kappa_z$), however the difference 
in the transverse trap stiffness ($\kappa_x$) is far more 
noticeable. The same dimer was trapped at each of the axial 
equilibrium positions and the transverse force was evaluated. 
While in all three cases the dimer can be trapped the linear
range where that would typically associated with a stable 
trap is far narrower in the 'standard' orientation compared 
to the 'inverted' cases. This highlights one of the challenges 
involved with studying asymmetric particles, even though its 
a simple enough process to trap them they maybe characterised 
very differently depending on their relative position and 
orientation towards the focus. This can have a significant 
impact on rheological studies - or attempting to probe any 
local property - as the variance in trap strength can result 
in large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's 
		centre of mass spheres, where a positive force 
		indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation 
		parameters are used here as in fig~\ref{fig:paradigmatic}
		(a) and (c). The blue curve representing the force 
		response for a dimer in its standard orientation, orange 
		being the inverted case, and green the same case but 
		placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

\newpage
For completeness the harmonic traps were located for dimers
across a range of size ratios - from $a_{I}/a_{II} = 1$ to 
$a_{I}/a_{II}=10$ - while also recording the trap stiffness 
for each trap. The same simulation parameters are used here 
as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease 
the dimer begins to approximate a single homogenous sphere 
- at least in terms of location and trap strength. However, 
for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
to $a_{I}/a_{II}=4$), a second equilibrium position is found 
below the trapping focus. Previous work using the ray-optics 
model have confirmed even in the case that two spheres begin 
separated the electric field will align the particles as such 
that they make contact and are trapped together about a single 
trapping position \cite{Xu2005}. Furthermore it has been shown 
through proper manipulation of the Gaussian or Bessel beam 
modes that any number of trapping potentials can be developed 
\cite{Shahabadi2020} for nanoparticles. This result however, 
is the first example of an orientation dependent trapping 
situation using only a $TEM_00$ beam. Typical experimental 
arrangements cannot determine much information on the axial 
position of a trapped particle relative to the trap focus; 
this result indicates not only that dimers can be trapped 
in multiple axial positions but also their trapping behaviour 
is heavily dependent on said axial position. As such it is 
necessary that positional information in the z-axis can be 
elucidated if multiple spheres are trapped simultaneously. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers with varying 
		size ratio, dashed lines represent unstable traps whereas solid lines 
		are for stable equilibrium positions. (a) shows that dimers while in their
		'standard' orientation will always have a single equilibrium position. 
		(b) shows that when the same dimer is in its' 'inverted' orientation 
		can be trapped in two axial positions, one below the focus and one above 
		the focus.}
	\label{fig:eq_positions}
\end{figure}

\newpage
\subsection{Non-trivial equilibrium configurations}
\label{sec:off-axis}
Computing the equilibrium positions when a dimer is aligned 
with the electric field is relatively simple as the orientational 
torque is minimised (see Eq.\ref{eq:opt_torque}). Meaning once 
trapped the dimer is unlikely to change orientation enough to 
escape the trap. Regardless, that does not rule out the 
possibility that there is a stable configuration where the 
orientation not strictly vertical, in fact most experimental 
work with symmetric nano-dimers will trap them lying perpendicular 
to the beam direction \cite{Ahn2018, Reimann2018}. Unlike in Sec.~
\ref{sec:eq_positions} we cannot simply measure the optical force 
and torque as the parameter space is too large and determining if 
a particular position and orientation is stable is not clear based 
solely on force and torque measurements \cite{Bui2017}. Using the 
same simulation parameters as before we ran a number short 
simulations (total simulation time was $0.005$ s) with the laser 
power increased to $500\ mW$. Each simulation started with the 
dimer in a different starting position and orientation, due to the 
high laser power the dimers either escaped the trap or were stably 
trapped. The $z-\theta$ phase space - where $\theta$ is the angle 
between the direction of beam propagation and the dimer's orientation 
vector ($\theta=0^\circ$ is the 'standard' orientation) - can be 
divided into different regions depending on which equilibrium 
configuration is reached.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{0pt}[0pt][0pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
		\caption{}
	\end{subfigure}
	%\captionsetup{hangindent=2.35cm}
	\caption{Map of $z-\theta$ phase space using a dimer of
		size ratio 2 with a laser power of 500 mW ($\theta=0
		\circ$ is the 'standard' orientation and $\theta=180
		^\circ$ is the 'inverted' orientation). The stable 
		configurations are indicated by the larger circles 
		and the starting conditions are colour coded to match 
		the stable point they end up in. Right hand render 
		shows a dimer in its off-axis configuration.}
	\label{fig:off_axis}
\end{figure}

Interestingly while the trap strength of these off-axis traps 
are similar in magnitude to the vertically aligned dimers, 
but when the laser power is lowered (around $5\ mW$) the traps 
become metastable; after reaching its equilibrium configuration 
the particle behaves similarly to a typically trapped dimer but 
the trapping potential is small enough that the dimer escapes 
in as little time as less than a second or after nearly a full 
3 seconds. Running similar simulations but for vertical 
configurations sees the dimer remaining trapped, even after 30 
seconds of run time, indicating that the trapping potential is 
far greater than the thermal energy. This suggests that the 
reason this off-axis configuration is metastable is due to the 
increased rotational freedom. Similar configurations have been 
explored with ellipsoids; Zhu \textit{et al.} looked at the 
dynamics of various elliptical particles and found that regardless 
of shape or initial orientation the particle would tend towards
either a purely vertical or purely horizontal orientation 
\cite{Zhu2021}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Continuous rotational motion in circularly polarised light}
One aspect that has yet to be covered in depth with regards 
to spherical aggregates of any construction is their interaction 
with circularly polarised light. For homogenous spheres the 
optical torque is regarded as being negligible as the spin 
density cannot impart angular momentum while propagating in a 
homogenous medium. Dimers however, have been shown to experience 
an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
while trapped in a circular polarised beam. In our simulations 
we found that dimers would rotate about their long axis when 
trapped in circularly polarised light. In this section we want 
to discuss how this behaviour is influenced by size, position, 
and orientation; and furthermore, we wish to address possible 
explanations for this behaviour, as none of the current
theories into optically induced rotation seem plausible.

@@

To that end, we simulated the motion of an optically trapped dimer in 
beams of varying polarisation ($NA=1.2$, $P=100\ mW$), the dimer is 
composed of polystyrene ($n_p=1.59$, $n_m=1.33$). Each simulation was 
run for $1$ second ($\Delta t =10^-5$) and at the end we looked at the 
orientational time series; the dimer's orientation is recorded as a 
quaternion which can be easily converted to a 3-dimensional rotation 
matrix. By considering only the transverse components ($U_{1,x}$, 
$U_{1,y}$, $U_{2,x}$, \& $U_{2,y}$) of the rotation matrix and taking 
the Fourier transformation of their time series reveals the rotational 
frequency. The laser power is set to $100\ mW$ to avoid large thermal 
fluctuations and so that the Fourier series of the transverse components 
approximates $\delta(\omega_{rot}-f)$ - the Dirac delta function centred 
at the rotational frequency $\omega_{rot}$.

>>>>

\chapter{Complex Langevin dynamics of spherical dimers}
\label{chapter:langevin_dynamics}
Much of the calibration theory discussed in Chapter 2 
assumes that the target particle in question is a single 
sphere, one whose scattering and motion is easily computed. 
However, while working with dense colloidal suspensions, 
one often ends up trapping more than one sphere. Li and 
Arlt \cite{Li2008} studied the case of two microspheres 
trapped in a single OT and found that multiple trapped 
beads could be mistaken for a single trapped bead with an 
altered trap stiffness (see \ref{sec:harmonic_traps}). 
Theoretical studies on the case of two trapped microspheres 
by Xu~\textit{et~al.} \cite{Xu2005} employed a ray-optics 
based model to show that the two trapped beads are brought 
into physical contact with each other by optical forces 
and they also calculated the axial equilibrium positions 
of the two trapped beads as a function of their size. 
Experiments in \cite{Praveen2016} confirmed that the two 
trapped beads indeed experience different trap stiffnesses 
in the vicinity of the same potential well.

There are further discussions looking into the dynamics of 
a whole host of asymmetrically shaped particles (e.g. 
ellipsoids, cylinders, amorphous solids, disks, etc) 
\cite{Loudet2014, ShengHua2005, Chetana2022}, their results 
all showing that predicting the behaviour an arbitrary shaped 
particle comes with great difficulty due to the fact that the 
optical force is dependent on a greater number of variables 
such as orientation and size factors.

As a paradigmatic example, we consider a dimer suspended in 
water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
a Gaussian beam (more specifically a Laguerre-Gaussian 
beam of mode $[0,0]$, see \eqref{eq:incident}), the beam 
is focused by a objective with numerical aperture of 1.2 
and is x polarised. The size ratio of the two spheres is 
given by $a_{I}/a_{II} = 2$ where $a_{I}$ is kept at $1\ 
\mu m$ unless specified otherwise; the dimer's orientation 
is given by a unit vector connecting the centres of both 
spheres (see \ref{sec:sim_parameters}). We define the 
'standard' orientation as being aligned with the direction 
of beam propagation direction - and therefore the 'inverted' 
orientation is defined when the dimer is orientated against 
the direction of beam orientation (see left hand side of 
figure~\ref{fig:paradigmatic}).

With the initial goal of the project being to induce 
nucleation events via a rotating sphere, the aim of this 
chapter is to - in a limited capacity simulate and investigate 
the influence of a second particle being bound to our target 
sphere. The choice of a dimer, instead of an amorphous solid 
that might better represent a growing crystalline solid, 
allows us to consider how the dynamics of the aggregate change 
by varying a single parameter, namely the size ratio of the 
two spheres. Additionally, attempting to simulate an amorphous 
aggregate is rather difficult as calculating the optical force 
and torque is computationally slow.

We utilise a python library that utilises \textit{mstm} and
\textit{ott} to simulate the dynamical behaviour of spherical
aggregates. A dimer is defined by its refractive index and the 
size ratio of the two spheres.  By specifying the optical 
properties of the target \textit{mstm} generates the targets 
T-matrix which can then be used by \textit{ott} to compute the 
optical forces generated by a focused beam.

In this chapter we will consider how the addition of a 
second sphere changes the trapping dynamics by introducing 
multiple equilibrium positions. Furthermore, we look at how 
dimers interact with circularly polarised light, this is 
especially pertinent as the original plan was to utilise 
circularly polarised light to generate fluid flow in a 
supersaturated solution. And lastly, we demonstrate how a 
quadrant photo diode, performs in characterising the actual 
interactions between a dimer and the optical trap. We do so 
to demonstrate that much of the dimer's trajectory information 
is either lost or poorly described by simple calibration 
techniques, making their motion difficult to characterise.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Simulation Details}
As a paradigmatic example, consider a dimer suspended in 
water ($n_p = 1.59, n_m = 1.33$) located at the focus of 
a Gaussian beam (more specifically a Laguerre-Gaussian 
beam of mode $[0,0]$, see \eqref{eq:incident}). The beam 
is focused by a objective with numerical aperture of 1.2.
The polarisation is defined using the Jones vector (see 
section \ref{sec:rot_pol} for further details) unless 
stated otherwise the Jones vector is kept at $[1.0, 0.0i]$. 
The size ratio of the two spheres is given by $a_{I}/
a_{II} = 2$ where $a_{I}$ is kept at $1\ \mu m$ unless 
specified otherwise. The beam focus is used as the 
origin of the coordinate system, with positive z going 
in the direction of beam propagation. For simulations 
we assume that the dimer is suspended in water at room 
temperature ($T = 298K,\ \eta = 1.0x10^{-3}Pa\ s$).

In order to compute the magnitude of the Brownian forces we
need to know the dimer's diffusion tensor \textbf{D}. Unlike
for spheres where the translational and rotational diffusion 
coefficients are equal along all three Cartesian axis', a
dimer's translational and rotational diffusion is dependent
on whether we consider the its long or short axis. We use
the results from \cite{Nir1973} to compute the dimensionless 
diffusion tensor and then scale it according to the radius of
the largest sphere ($a_I$). The dimer's centre of diffusion
is located a distance of $a_I\zeta$ from the point of contact
between the two spheres, the coefficient $\zeta$ is based on
the size ratio of the two spheres. We use the centre of 
diffusion as the target origin that is provided to \textit{mstm}.

The dimer's orientation is given by a rotation vector in 
the form: 
\begin{align}
	\begin{pmatrix}
		U_{1,x} & U_{2,x} & U_{3,x} \\
		U_{1,y} & U_{2,y} & U_{3,y} \\
		U_{1,z} & U_{2,z} & U_{3,z}
	\end{pmatrix}
\end{align}
Where $U_3$ represents the orientation vector that passes
through the centres of both sphere's (see \ref{sec:sim} 
for a full breakdown), $U_1$, and $U_2$ are orthogonal to 
$U_3$ but have no physical representation for a dimer. We 
can represent $U_3$ by it's spherical coordinates ($\theta$ 
and $\phi$), where $\theta$ is the angle between the 
orientation vector and the direction of propagation and 
$\phi$ as the angle between the orientation vector and 
the x-axis. We define the 'standard' orientation as 
being aligned with the direction of beam propagation 
direction ($\theta=0^\circ$) - and therefore the 'inverted' 
orientation is defined when the dimer is orientated against 
the direction of beam orientation ($\theta=180^\circ$) - see
fig~\ref{fig:paradigmatic} for a visual of the laboratory and
particle frame.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Positional and Orientational dependence of Trapping forces}
\label{sec:eq_positions}
If we wanted to start from first principles and determine 
the trap strength on our target particle the first step 
would be to locate the equilibrium position relative to 
the trap focus. For a single sphere it is easy to enough 
to understand that its centre of mass will be drawn to 
focal point of the laser due to gradient forces, once there 
the force is analogous to a harmonic spring with a fixed 
trap stiffness (see fig.\ref{fig:harmonic_trap}). Now, if 
we consider instead a dimer, we now have two spheres both 
being drawn to the focus along by the same gradient force; 
in addition the scattering force is significantly more complex 
due to both spheres scattering the electromagnetic fields. 
This mutual scattering between individual spheres is what 
makes simulating spherical aggregates far more difficult 
compared to a single sphere, and even harder still to 
predict the position where the dimer's centre of mass is 
in equilibrium.

@@

This is different to a single sphere, namely in that a
single sphere should only have one equilibrium position
and that it is much closer to the trap focus ($\approx
0.05\mu\ m$ from the focus). We can see that both 
equilibrium positions have comparable axial trap 
stiffness ($\kappa_z$), however the difference in the 
transverse trap stiffness ($\kappa_x$) is far more 
noticeable. The same dimer was trapped at each of the 
axial equilibrium positions and the transverse force was 
evaluated. While in all three cases the dimer can be 
trapped the linear range where that would typically 
associated with a stable trap is far narrower in the 
'standard' orientation compared to the 'inverted' cases. 
This highlights one of the challenges involved with 
studying asymmetric particles, even though its a 
simple enough process to trap them they maybe 
characterised very differently depending on their 
relative position and orientation towards the focus. 
This can have a significant impact on rheological 
studies - or attempting to probe any local property 
- as the variance in trap strength can result in 
large errors over repeated measurements. 
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{transverse_force.png}
	\caption{Plots of force vs displacement of the dimer's 
		centre of mass spheres, where a positive force 
		indicates the dimer is directed right on the x-axis, 
		and vice versa for a negative force. The same simulation 
		parameters are used here as in fig~\ref{fig:paradigmatic}
		(a) and (c). The blue curve representing the force 
		response for a dimer in its standard orientation, orange 
		being the inverted case, and green the same case but 
		placed below the focus.}
	\label{fig:transverse_force}
\end{figure}

\newpage
For completeness the harmonic traps were located for dimers
across a range of size ratios - from $a_{I}/a_{II} = 1$ to 
$a_{I}/a_{II}=10$ - while also recording the trap stiffness 
for each trap. The same simulation parameters are used here 
as for figures \ref{fig:paradigmatic} \& \ref{fig:transverse_force}. 
As shown in Fig.~\ref{fig:eq_positions} $a_{II}$ decrease 
the dimer begins to approximate a single homogenous sphere 
- at least in terms of location and trap strength. However, 
for intermediate sized dimers (between $a_{I}/a_{II} = 1.1$ 
to $a_{I}/a_{II}=4$), a second equilibrium position is found 
below the trapping focus. We should note that there are also
unstable traps that exist, these are also depicted in 
fig.~\ref{fig:eq_positions}. While the optical force
is zero at these positions the gradient $\delta F_z/\delta z$
is positive and so any Brownian motion would displace the 
dimer, either ejecting it from the optical trap or letting
it be drawn towards another harmonic trap. 
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions.png}
		\caption{}
		\label{eq_pos}
	\end{subfigure}\hfill % <-- "\hfill"
	\medskip
	\begin{subfigure}{0.85\linewidth}
		\includegraphics[width=\linewidth]{Equillibrium_positions_inverted.png}
		\caption{}
		\label{eq_pos_inverted}
	\end{subfigure}
	\caption{Equilibrium positions of optically trapped dimers 
		with varying size ratio, dashed lines represent unstable 
		traps whereas solid lines are for stable equilibrium 
		positions. (a) shows that dimers while in their 
		'standard' orientation will always have a single 
		equilibrium position. (b) shows that when the same dimer 
		is in its' 'inverted' orientation can be trapped in two 
		axial positions, one below the focus and one above the 
		focus.}
	\label{fig:eq_positions}
\end{figure}

Previous work using the ray-optics model have confirmed even 
in the case that two spheres begin separated the electric field 
will align the particles as such that they make contact and 
are trapped together about a single trapping position 
\cite{Xu2005}. Furthermore it has been shown through proper 
manipulation of the Gaussian or Bessel beam modes that any 
number of trapping potentials can be developed 
\cite{Shahabadi2020} for nanoparticles. This result however, 
is the first example of an orientation dependent trapping 
situation using only a $TEM_00$ beam. Typical experimental 
arrangements cannot determine much information on the axial 
position of a trapped particle relative to the trap focus; 
this result indicates not only that dimers can be trapped 
in multiple axial positions but also their trapping behaviour 
is heavily dependent on said axial position. As such it is 
necessary that positional information in the z-axis can be 
elucidated if multiple spheres are trapped simultaneously.

\newpage
\section{Non-trivial equilibrium configurations}
\label{sec:off-axis}
As mentioned previously in \ref{sec:opt_torque} the 
torque applied to an anisotropic scatterer can be 
broken down into two categories: torque generated 
from the field polarisation and torque generated 
by the gradient force of the electric field. The 
former is generally only applicable for objects 
that are birefringent whereas the latter is 
applicable to dimers due to their elongated shape.
Since both spheres are attracted towards the centre
of intensity via the gradient force a dimer will 
experience a restoring torque that keeps the it
in an equilibrium orientation. For any elongated 
particle the torque is minimised when the long 
axis is aligned either parallel or perpendicular
with the direction of the electric field. In this 
section we try to find equilibrium configurations 
(positions and orientations where both the optical
force and torque are minimised).

Computing the equilibrium positions when a dimer is 
aligned with the electric field is relatively simple 
as the orientational torque is minimised (see Eq.
\ref{eq:opt_torque}). Meaning once trapped the 
dimer is unlikely to change orientation enough to 
escape the trap. Regardless, that does not rule out 
the possibility that there is a stable configuration 
where the orientation not strictly vertical, in fact 
most experimental work with symmetric nano-dimers will 
trap them lying perpendicular to the beam direction 
\cite{Ahn2018, Reimann2018}. Unlike in Sec.~
\ref{sec:eq_positions} we cannot simply compute the 
optical force and torque as the parameter space is too 
large and determining if a particular position and 
orientation is stable is not clear based solely on 
force and torque measurements \cite{Bui2017}. Using 
the same simulation parameters as before we ran a 
number short simulations (total simulation time was 
$0.005$ s) with the laser power increased to $500\ mW$. 
Each simulation started with the dimer in a different 
starting position and orientation, due to the high 
laser power the dimers either escaped the trap or 
were stably trapped. We chose to only consider the $z-
\theta$ phase space in order to simplify the explanation,
considering all possible parameters would increase the 
simulation time significantly. The $z-\theta$ phase space 
can be divided into different regions depending on which 
equilibrium configuration is reached.
\begin{figure}[h!]
	\centering
	\begin{subfigure}{0.67\linewidth}
		\includegraphics[width=\linewidth]{off_axis_trap.png}
	\end{subfigure}
	\begin{subfigure}{0.32\linewidth}
		\raisebox{0pt}[0pt][-10pt]{\makebox{}\includegraphics[width=\linewidth, keepaspectratio]{off_axis_render.png}}
	\end{subfigure}
	%\captionsetup{hangindent=2.35cm}
	\caption{Map of $z-\theta$ phase space using a dimer 
		of size ratio 2 with a laser power of 500 mW 
		($\theta=0^\circ$ is the 'standard' orientation 
		and $\theta=180^\circ$ is the 'inverted' 
		orientation). The stable configurations are 
		indicated by the larger circles and the 
		starting conditions are colour coded to match 
		the stable point they end up in. Right hand 
		render shows a dimer in its off-axis configuration.}
	\label{fig:off_axis}
\end{figure}

Where the each starting location is colour coded to
match the final equilibrium position reached, empty
regions indicate that the trap was escaped. As 
expected we see that the dimer has three separate 
traps when the dimer is close to its vertical 
orientation, similar to what we see from figs.~
\ref{fig:paradigmatic} and \ref{fig:eq_positions}. 
When the beginning orientation is close to horizontal 
there exists a $4^th$ harmonic trap. Each of these 
traps, has its own 'basin' of attraction around 
which the dimer is drawn into, the strange shape 
of these basins is likely due to how the asymmetric 
shape interacts with the optical trap.

Interestingly while the trap strength of these off-axis 
configurations are similar in magnitude to the vertically 
aligned dimers, but when the laser power is lowered 
(around $5\ mW$) the traps appear metastable. With the 
vertical configurations even after $30\ s$ of simulation
time the dimer is still trapped whereas the off-axis 
configuration can be escaped after only a few seconds of
simulation time. The suggested that the configuration is 
metastable is due to the increased rotational freedom. 
Similar configurations have been explored with ellipsoids; 
Zhu \textit{et al.} looked at the dynamics of various 
elliptical particles and found that regardless of shape or 
initial orientation the particle would tend towards either 
a purely vertical or purely horizontal orientation 
\cite{Zhu2021}. However, in our case the dimer never returns
to a vertical orientation, suggesting their is a potential 
barrier separating the different configurations. In order
to build a better understanding of the potential landscape 
we need to consider how the force and torque vary for a 
dimer in the $z-\theta$

\subsection{Force-Torque landscape of an arbitrary dimer}
We use \textit{ott} to compute the optical force 
and torque for the same dimer as used in fig.~
\ref{fig:off_axis} in different positions or 
orientations, this is depicted in fig.\ref{fig:force_torque}. 
The magnitude of the force and torque are indicated
by the colour intensity, with white indicating either
zero force or torque. This helps show where a
dimer may end up in equilibrium by finding 
where both the torque and force go to 0.
\begin{figure}[h!]
	\centering
	\includegraphics[width=\linewidth]{force_torque.png}
	\caption{Force-Torque landscape over the $z-\theta$ 
		phase space with a dimer is a size ratio of 2, 
		laser power of 100mW. Left: 2D plot of the 
		optical force in the axial direction as a 
		function of position and orientation. Right: 
		2D plot of the optical torque about the x-axis 
		as a function of position and orientation.}
	\label{fig:force_torque}
\end{figure}

The force and torque landscape provides some insight 
into how the dimer can inhabit an off axis equilibrium
orientation. Firstly, lets consider a typical trap, 
while the dimer is close to vertical; in all of these 
cases the optical force and torque directs the dimer 
back to the configuration. In this case the potential 
well is positive whether the dimer moves along the 
z-axis or is rotated. We can assume that in both cases
the translational and rotational stiffnesses are close
enough to be modelled as harmonic potentials.

However when in an off-axis configuration ($z\approx0$ 
\& $90^\circ\le\theta\le115^\circ$) the optical torque 
is working against the optical force, trying to move it 
out of equilibrium while the optical force returns it to 
equilibrium. This means that the potential well is more 
akin to a saddle point, the only reason that the optical
trap is not escaped immediately is because the the 
translational trap stiffness is significantly greater than
the gradient of the optical torque ($\kappa_z\gg\kappa_
\theta$). Therefore it is more accurate to say that the 
trap is not metastable, as when the trap is escaped it 
is not because the Brownian displacement is greater than
the potential well, rather the dimer is randomly 
traversing along the x-axis (aka random Brownian 
rotation) and has escaped the translational potential 
well.

The force-torque landscape also provides an explanation
to the boundaries between the different trapping 'basins'
in fig.~\ref{fig:off_axis}. The first point of interest 
is the 'dead zone' around the orange basin. Looking at 
fig.~\ref{fig:force_torque} we can see that around that
region ($z \approx 2.5\mu$ \& $\theta\approx 110^\circ$)
the optical torque changes sign as you move closer to 
the trap focus. There is no clear physical reason why
the optical torque change direction so abruptly but a 
consequence of it is that dimers that are too close to 
this region of the torque landscape are ejected from 
the optical trap as the are drawn down towards the trap
focus.

The only pellicular result is the observation that the 
blue basin can still be reached well below the trap 
focus. As shown by fig.~\ref{fig:off_axis} there are a 
handful of starting configurations that end up in the 
upper trap while in the inverted orientation. Even 
consulting fig.~\ref{fig:force_torque} does not indicate
any reason why this occurs. The only conclusion we can 
draw is that the orange and green basins of attraction 
are small enough that it is possible for a dimer to 
avoid both. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Continuous rotational motion in circularly polarised light}
One aspect that has yet to be covered in depth with regards 
to spherical aggregates of any construction is their interaction 
with circularly polarised light. For homogenous spheres the 
optical torque is regarded as being negligible as the spin 
density cannot impart angular momentum while propagating in a 
homogenous medium. Dimers however, have been shown to experience 
an optical torque \cite{Vigilante2020, Ahn2018, Reimann2018} 
while trapped in a circular polarised beam. In our simulations 
we found that dimers would rotate about their long axis when 
trapped in circularly polarised light. In this section we want 
to discuss how this behaviour is influenced by size, position, 
and orientation; and furthermore, we wish to address possible 
explanations for this behaviour, as none of the current
theories into optically induced rotation seem plausible.

@@

To that end, we simulated the motion of an optically 
trapped dimer in beams of varying polarisation ($NA=1.2$, 
$P=100\ mW$). As mentioned previously the beam polarisation
is given using the Jones vector which describes the magnitude
of the x and y components of the electric field (so the vector
$[1.0, 0.0i]$ describes an electric field that only oscillates
in the x direction). For a linearly polarised beam, both 
components are in phase with one another, for an elliptical or
circularly polarised beam we adjust the phase of each component. 
So a Jones vector of $[1.0, 1.0i]$ describes a beam where the
y component is $\pi/2$ radians out of phase with the x-component 
resulting in an electric field that rotates clockwise about
its direction of propagation. For each simulation we vary
the phase in increments of $\pi/24$.
